---
output:
  html_document: default
  pdf_document: default
---

# Proportions and Percents

We can imagine lots of lists do not contain numbers. 
Instead of number, these lists contain not-numbers.
We sometimes refer to not-numbers as "qualitative" (as opposed to "quantitative") values. 
We might have a list that contains the class standing of each student in our class. 
Rather than numbers, this list of not-numbers contains qualitative values like "Freshman," "Sophomore," "Junior," and "Senior." 

Similarly, at the end of the semester, I have two lists of grades. 
The first lists numbers that represent each student's points in the class.
This list of quantitative values ranges from 0 to 100.
But you care more about the second list, which contains each student's letter grade in the class. 
This list of not-numbers contains the qualitative values A, B, C, D, and F. 

**Exercise:** Similarly, we might have lists that contains partisanships, ideologies, regime types, and so on. What qualitative values might we find in a list of partisanships? Ideologies? Regime types?

When we deal with lists that contain qualitative values, **we cannot use the average** because we cannot sum the not-numbers. 

Instead of an averarge, we use *proportions*. 
We choose a particular qualitative values and compute the proportion of the list that falls has that value. 
Note that this "particular category" might contain several values, such as freshman or sophomore (underclassmen) or A, B, or C (passing grades). 
For example, we say a certain proportion of the class are freshman. 
Or we say that a certain proportion of a class received an A, B, or C.

## Calculating Proportions

To calucate a proportion, simply count the number of entries in the list that fall in the particular category and divide that number by the total number of entries.

$$\text{proportion in the category} = \dfrac{\text{number of entries that fall into the category}}{\text{total number of entries}}$$

## Converting Proportions to Percents

Proportions serve an important purpose---they make math easy. 
However, the scale makes proportions difficult to write or talk about.
To make the math easy, we use proportions. To make the interpretation easy, we use percents.

To convert a proportion to a percent, we simply multiply the proportion times 100%.

$$\text{percent in the category} = \text{proportion in the category} \times 100\%$$.

As a general rule of thumb, I recommend that you convert proportions to percents, but only *after you complete all calculations.*^[Sometimes, calculation that work for proportions also work for percents. However, sometimes they do not. Because of these, I recommend doing all calculations with proportion and convert the final answer into a percent.]

**Example**: Compute the percent of sophomores in the class from the data in Table \@ref(tab:grades). 

```{r grades, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(babynames)
data(babynames)

class <- c("Freshman", "Sophomore", "Junior", "Senior")
grade <- c("A", "B", "C", "D", "F")
name <- sample(babynames$name, size = 15, prob = babynames$prop)

set.seed(12345)
df <- data.frame(Name = name, 
                 Class = sample(class, size = 15, replace = TRUE), 
                 Grade = sample(grade, size = 15, replace = TRUE, prob = c(1, 2, 3, 1, 1)))
kable(df, booktabs = TRUE, align = "c",
      caption = "Class standing and letter grades of 15 hypothetical students.") %>% 
  kable_styling(position = "center", latex_options = c("striped", "hold_position"))
```

> First count the number of sophomores in the list. We have `r sum(df$Class == "Sophomore")`. Next, count the number of total students in the class. We have `r length(df$Class)`. Finally, divide the number of sophomores by total number of students. We have $\frac{`r sum(df$Class == "Sophomore")`}{`r length(df$Class)`} = `r sum(df$Class == "Sophomore")/length(df$Class)`$.

> To convert the proportion 0.2 to a percent, we simply multiply it by 100%. This gives $0.2 \times 100\% = 20\%$. We can then say "20% of the students in the class are sophomores."

**Exercises**

1. Compute the percent of students in the class that are sophomores. Compute the percent that are juniors or seniors.
1. Compute the percent of students that received an F. Compute the percent that received an A or a B.
1. Compute the percent of seniors that received an A. Compute the percent of freshmen that received an A.

## The Mathematics of Proportions

The approach I describe above to compute a proportion works perfectly. 
However, it does not clearly connect to the past (or future) discussion of averages. 

If we can connect proportions and averages, then we reduce our work later.
By thinking of a proportion as a special kind of average, we can apply the same ideas to two concepts, cutting the ideas we need to learn and understand in half.

To see that a proportion is a special type of average, we can simply convert the list of not-numbers into a list of numbers. 
To create the new list of numbers, assign the number 1 to the entries that fall into the particular category and assign the number 0 to the other entries. 
This creates a new list of numbers that contains only 0s and 1s. 
I call this type of list a 0-1 list. 
The proportion is simply the average of this 0-1 list. 

**Important: We can think of a proportion as an average of a 0-1 list.**

Example: Using a 0-1 list, compute the proportion of students in Table \@ref(tab:grades) that are sophomores.

> Create a 0-1 list where 1 indicates sophomores and 0 indicates. Table \@ref(tab:gradesw01list) includes this 0-1 list. Now simply average the 0-1 list. We have $\frac{\text{sum of 0-1 list}}{\text{number of entries in 0-1 list}} = \frac{`r sum(df$Class == "Sophomore")`}{`r length(df$Class)`} = `r sum(df$Class == "Sophomore")/length(df$Class)`$.

```{r gradesw01list, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(babynames)
data(babynames)

class <- c("Freshman", "Sophomore", "Junior", "Senior")
grade <- c("A", "B", "C", "D", "F")
name <- sample(babynames$name, size = 15, prob = babynames$prop)

set.seed(12345)
df <- data.frame(Name = name, 
                 Class = sample(class, size = 15, replace = TRUE), 
                 Grade = sample(grade, size = 15, replace = TRUE, prob = c(1, 2, 3, 1, 1))) %>%
  mutate(`0-1 List[note]` = ifelse(Class == "Sophomore", 1, 0))

kable(df, booktabs = TRUE, align = "c",
      caption = "Class standing and letter grades of 15 hypothetical students.") %>% 
  kable_styling(position = "center", latex_options = c("striped", "hold_position")) %>%
  add_footnote(c("1 indicates sophomores and 0 indicates not-sophomores."),
               notation = "symbol")
```

**Exercise:** Suppose I want to evaluate the graduate admissions process at a university, so I put together a list containing the gender of the admitted graduate students and obtain M, M, F, F, M, M, M, M, F, M, and M.
Try two ways to calculating the proportion of women admitted. 
First, use inital method I described: count the number of women in the list and divide that by the total number of entries. 
Next, use the more mathematical way I described: write down the implied 0-1 list (replace F with 1 and everything else with 0) and then average the list. 
Are these two approaches different or the same? 
Can you explain why the results are identical?

## SD of a 0-1 List

If we convert a list of not-numbers into a 0-1 list, then we can compute the SD of the new list. 
Just like with other lists of numbers, the SD of a 0-1 list is the R.M.S. of the deviations from average. 
We can compute the SD the usual way. 
But, as you well know, the usual way takes time. 
If we have a 0-1 list, though, we can use a shortcut.

$$\text{SD of a 0-1 list} = \sqrt{\text{ave. of list} \times (1 - \text{ave. of list})}$$

Notice that the average of a 0-1 list (the proportion) determines the SD *exactly*. 
If the average of a 0-1 list is 0.25, then the SD is exactly $\sqrt{0.2 \times 0.8} = 0.4$. 
If the average of a 0-1 list is 0.5, then the SD is exactly $\sqrt{0.5 \times 0.5} = 0.5$. 

Because of this, we usually pay little attention to variation when working with lists of qualitative values---the proportion (or percent) contains all the information we need.

However, this shortcut will help us later, so remember it.

Exercise: Suppose a 0-1 list that contains 6 0s and 4 1s. What is average of this list? What proportion are 1s? What is the SD of the list?

## Two Facts about Proportions and Percents

We can say two things about proportions and percents: 

1. A proportion is at least 0 and at most 1. Similarly, the percent is at least 0% and at most 100%. If none of the entries fall into the category, then the top (numerator) of the formula for a proportion equals 0, and therefore the whole proportion equals 0. If all of the entries fall into the category, then the top of the formula for the proportion proportion (numerator) equals the bottom of the proportion (denominator), and thefore the whole proportion equals 1. 
2. The proportion that do not fall into a particular category is 1 minus the proportion that fall into that category. It turns out that the percent that do not fall into a particular category is 100% minus the percent that fall into that category.

**Exercises**

1. In a Gallup poll conducted in late December of 2017, 39% of respondents said they "approve" of the job that Donald Trump is doing as president. What percent did not say they approve? In the same survey, 55% said they "disapprove" of the job that Donald Trump is doing as president. How is this possible?
1. I wrote a computer program to analyze a data set. I found that 120% of citizens living in a district received contact from a political campaign. Did I do something wrong?

## Review Exercises

1. Table \@ref(tab:govtcontroldata) shows the partisan control of each branch of the 49 U.S. states (excluding Nebraska) in 2011. Compute the proportion of states with a Republican governor. Repeat for house and senate. Is the proportion of Democratic governors, houses, and senates necessarily $1 - \text{proportion Republican}$? Why or why not? Which party has most control in the U.S. states? Can you think of any ways that control of state governments affect national politics? 

    ```{r govtcontroldata, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
    # load packages
    library(tidyverse)
    library(magrittr)
    library(kableExtra)
    
    # load and clean data for example
    df <- read_rds("data/state-governments.rds") %>%
      filter(year == 2011) %>%
      select(-year) %>%
      arrange(state)
    
    # print md table
    knitr::kable(df, align = "c", caption = "The partisan control of each branch of the 49 (excluding Nebraska) state governments in 2011.") %>%
            kable_styling(full_width = FALSE, latex_options = c("striped", "hold_position"), font_size = 9)
    ```

1. Obtain a copy of Gerber, Green, and Larimer's 2008 article "Social Pressure and Voter Turnout: Evidence from a Large-scale Field Experiment" published in the *American Political Science Review* (Volume 102, Issue 1, pp. 33-48). Complete the following tasks:
    a. Read the introduction, pp. 33-34 though "Social Norms, The Calculus of Voting, and Prior Research." What question interests the authors? What type of question is it (normative, descriptive, causal)? Would you say that the question matters? Why?
    b. Read the section "Experimental Design," pp. 36-38 though "Results". Briefly describe the design of the study. When and where was the student conducted? Who was included in or excluded from the study? How did the researchers assign the subjects to the treatment and control groups?
    c. Examine the four mailers reproduced on pp. 43-46. Using your intuition about voter psychology and behavior, rank these mailers from most effective to least effective. Which mailer do you suspective makes the recipient most likely to vote? Least likely? Do you suspect any of the mailers have a negative effect (i.e., receiving the mailer makes the recipient less likely to vote than if she had received no mailer at all)?
    d. Table \@ref(tab:socialpressuredata) contains the data set `social-pressure`. Use these data to re-compute the percentages that the authors present in their Table 2 on p. 38. Fill these in the appropriate colum in Table \@ref(tab:socialpressuredata). 
    e. Estimate the average treatment effect by subtracting the proportion that voted in the control group from the proportion that voted in each treatment group (i.e., groups that received a mailer). Convert these changes in proportions to changes in percentages by multiplying by 100%.^[We must take care when discussing changes in percentages. A 10% increase in 50%, could mean either (a) $0.5 + 0.1 = 0.6 = 60\%$ or (b) $0.5 + (0.5 \times 0.1) = 0.5 + 0.05 = 0.55 = 55\%$. To make the change clear, we usually talk about changes in "percentage points," which clearly refers to (a). Thoughout these notes, I always describe changes in percents using percentage point changes (a).] According to these estimates, what treatment is most effective? Least effective? Do any treatments have a negative effect? Does the treatment effects match your guesses about the rankings?
    f. Comment on the ethics of this study? Would you describe this study as unethical? Why or why not?

    ```{r socialpressuredata, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
    # load packages
    library(tidyverse)
    library(magrittr)
    library(kableExtra)
    
    # load and clean data for example
    df <- read_rds("data/social-pressure.rds") %>%
      rename(`Condition` = condition,
             `Number of Subjects in Condition` = number_in_condition,
             `Number of Subjects in Condition that Voted` = number_that_voted) %>%
      mutate(`Percent of Subject in Condition that Voted` = "", 
             `Average Treatment Effect (in Percentage Points)` = ifelse(Condition == "Control", 
                                                                        "Not Applicable", ""))
    
    # print md table
    #knitr::kable(df, align = "c", caption = "The numbers of experimental subjects in each condition and the number of subjects in each condition that voted in Gerber, Green, and Larimer's (2008) experiment.")
    
    kable(df, booktabs = TRUE, align = "c", 
          caption = "The numbers of experimental subjects in each condition and the number of subjects in each condition that voted in Gerber, Green, and Larimer's (2008) experiment.") %>% 
      kable_styling(full_width = FALSE, latex_options = c("striped", "scale_down", "hold_position")) %>%
      column_spec(1) %>% 
      column_spec(2, width = "10em") %>% 
      column_spec(3, width = "10em") %>% 
      column_spec(4, width = "10em") %>% 
      column_spec(5, width = "10em")
    ```
    
1. In an article published in 2013 in the *American Journal of Political Science* (Volume 57, Issue 1, pp. 1-14) titled "Do Conditional Cash Transfers Affect Electoral Behavior? Evidence from a Randomized Experiment in Mexico," Ana L. De La O presents an argument that conditional cash transfer programs increase turnout. She summarizes her results (p. 11) as follows:

    > This article provides evidence on the electoral returns of the Mexican CCT by analyzing a unique randomized vari- ation in the duration of program benefits across eligible villages. The findings suggest that the targeted program led to an increase in voter turnout and incumbent vote shares. While previous work focuses on CCT persuasive effects, this article shows that the CCT pro-incumbent effects are mainly explained by a mobilizing mechanism.

    But what is CCT program? De La O summarizes Mexico's policy (p. 3): 
    
    > Responding to this context, the administration of President Ernesto Zedillo launched Progresa in 1997. The program consists of three complementary components: a cash transfer, primarily intended to subsidize food expenditure, delivered directly to the female heads of poor households; a scholarship, intended to compensate for the opportunity cost of child labor, thus enabling children to stay in school; and basic health care for all members of the household, with particular emphasis on preventive health care. All the components add up to an average transfer of US$35 per month, of which more than 85% is in cash. This is approximately 25% of the average poor, rural household income in the absence of the program. Transfers are paid every two months. Progresa is conditional because receipt of benefits is contingent upon school attendance, regular medical checkups, and attendance at platicas (meetings) related to health, hygiene, and nutrition issues.
    
    However, the government rolled out the policy in waves. Some locations---an "early immplementation" group---received the policy by late-1997. Other locations---a "late implementation" group---received the policy in early 2000. Importantly, each location was assigned to its group *randomly*.
    
    De La O was particularly interested in the effect of the policy on voter turnout. She writes that "assignment to early treatment leads to a 5 percentage point increase in turnout." To see the evidence for her claim, Table \@ref{tab:progressatab} shows the average turnout in each condition. This table shows that the average turnout across locations in the early condition was 68% and the average turnout across locations in the late condition was 63%. Based on these average turnout rates, it seems like the early implementation of the program increased turnout by about 5 percentage points on average.
    
    ```{r progressatab, echo=FALSE}
    # load packages
    library(tidyverse)
    library(rio)
    
    # load and clean data for example
    progresa_df <- import("data/progresa.rds")
    
    # find the average for each condition
    grouped_df <- group_by(progresa_df, condition)  # group the data frame
    smry_df <- summarize(grouped_df,  # find avg. for each group 
                         avg = mean(turnout, na.rm = TRUE)) %>%
      mutate(avg = paste0(round(100*avg), "%")) %>%
      rename(`Condition` = condition,
             `Average Turnout` = avg)
    kable(smry_df, booktabs = TRUE, align = "c", 
          caption = "The average turnout in each condition in De La O.'s study.") %>%
      kable_styling(latex_options = c("hold_position"))
    ```
    
    For a more detailed look, we can also examine histgrams showing the distribution of turnout across locations in the two conditions. Figure \@ref(fig:progresafig) shows these histograms. Does everything seem in order? Or can you spot something wrong? Do you think this likely undermines her conclusion, why or why not? (Perhaps you find Table \@ref{tab:progresatabmed} helpful.) 
    
    ```{r progressafig, echo=FALSE, fig.height=2, fig.width=5, message=FALSE, warning=FALSE, fig.align='center'}
    # plot histogram
    ggplot(progresa_df, aes(x = turnout, y = ..density..)) + 
      geom_histogram() +
      facet_wrap(~ condition) +
      scale_x_continuous(labels = scales::percent)  + # convert prop. to %
      labs(x = "Turnout", 
           y = "Density")
    ```
    
    ```{r progressatabmed, echo=FALSE}
    # load packages
    library(tidyverse)
    library(rio)
    library(kableExtra)
    
    # load and clean data for example
    progresa_df <- import("data/progresa.rds")
    
    # find the average for each condition
    grouped_df <- group_by(progresa_df, condition)  # group the data frame
    smry_df <- summarize(grouped_df,  # find avg. for each group 
                         med = median(turnout, na.rm = TRUE)) %>%
      mutate(med= paste0(round(100*med), "%")) %>%
      rename(`Condition` = condition,
             `Median Turnout` = med)
    
    kable(smry_df, booktabs = TRUE, align = "c", 
          caption = "The median turnout in each condition in De La O.'s study.") %>%
      kable_styling(latex_options = c("hold_position"))
    ```    