---
title: 'Codebook: `social-pressure` '
template: default
output:
  html_document: default
  pdf_document: default
---

```{r global_options, include=FALSE}
wd <- normalizePath("../../")
wd
knitr::opts_knit$set(fig.path='fig/',
               echo=FALSE, warning=FALSE, message=FALSE,
               root.dir=wd)
```

This data set is a cleaned version of the data from Gerber, Green, and Larimer's (2008) paper "Social Pressure and Voter Turnout" [[link](https://isps.yale.edu/sites/default/files/publication/2012/12/ISPS08-001.pdff)]. I obatined their replication data from the file `GerberGreenLarimer_APSR_2008_social_pressure.csv` [[link](https://isps.yale.edu/research/data/d001)] (see Excel.csv).

The following versions are available:

```{r, echo=FALSE}
list.files(path = "data/", pattern = "social-pressure.")
```


The data set is at the **condition level**, so that each row of the data set represents one of the five experimental conditions.

# Load Data

```{r, message=FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(rio)

# load data
sp_df <- import("data/social-pressure.rds")

# print entire data set
sp_df
```

# Variable Descriptions

## `condition`: Country Name

- Coding: The experimental condition.
- Type: factor

## `number_in_condition`: Number of Subjects in Each Condition

- Coding: The total number of experimental subjects in each condition.
- Type: integer

## `number_that_voted`: Number of Subjects that Voted

- Coding: The total number of experimental subjects in each condition *that voted*.
- Type: integer
