---
title: "Data Sources"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include=FALSE}
# load packages
library(tidyverse)
library(magrittr)
```

- `anes`
    - url: http://www.electionstudies.org/studypages/download/datacenter_all_datasets.php
    - accessed: 2018-01-02.
    - raw filename: `raw-data/anes_timeseries_2016.dta`.
    - raw codebooks: `raw-data/codebooks/anes_timeseries_2016_userguidecodebook.pdf` and `raw-data/codebooks/anes_timeseries_2016_varlist.pdf`.

```{r message=FALSE, warning=FALSE}
# load data
anes_df <- read_rds("anes.rds")

# quick look
glimpse(anes_df)
```


- `state-governments`
    - citation: Klarner, Carl, 2013, "State Partisan Balance Data, 1937 - 2011", hdl:1902.1/20403, Harvard Dataverse, V1. 
    - accessed: 2018-01-02. 
    - raw filename: `raw-data/Partisan_Balance_For_Use2011_06_09b.xlsx`.
    - raw codebook: `raw-data/codebooks/Partisan_Balance_For_Use2012_10_18_Codebook.docx`.


```{r message=FALSE, warning=FALSE}
# load data
state_governments_df <- read_rds("state-governments.rds")

# quick look
glimpse(state_governments_df)
```
